\subsection{Forward Propagation of Uncertainty}
\label{sec:forward}
The PC expansion created using an ensemble of 125 \geoclaw simulations
is now used as a surrogate to propagate prior input parameter uncertainty 
through the forward model.  Here we exploit the surrogate 
to study the statistics of the water surface elevation, and to conduct 
a global sensitivity analysis of the impact of the uncertain input parameters.

\subsubsection{Statistical analysis}
The mean of the QoI and its standard deviation is computed
from the PC coefficients as indicated in Equations~\eqref{eq:mean} and \eqref{eq:sigma}. 
In Figure~\ref{fig:ave} we plot the evolution of
the mean water surface elevation along with two standard deviations
bounds at the four gauges.  
Note that we only show the evolution after $t=6000~s$ where the uncertainty is significant,
reflected in a significant standard deviation as seen earlier from the realizations shown 
in Figure~\ref{fig:rlzs}. An interesting observation is that the
standard deviation in water surface elevations narrows at a few time instances
during the tsunami event.  The narrowing of the variance at these instances is
possibly due to the waves that arrive due to reflections from a single source
and then move away from the gauge location, and thus no variance in the water
surface elevation.

% or from a region where Manning's $n$ values have little influence over these
%waves.

The predicted mean water surface elevation  and its standard deviation 
can be used for comparison with the (DART) buoy observations collected during
the tsunami event. Figure~\ref{fig:compare} 
plots the PC mean water surface elevation with the observed 
water surface elevation at the different gauges locations
with variances indicated at few time instances. 
The plots show a good agreement between the PC simulations and the 
observations at different locations and at different times. 

The same statistical analysis can be performed for the
entire domain in 2D. Figure~\ref{fig:mean2d} (top row) shows
the PC mean water surface elevation for the considered computational
domain at three different times as indicated in the title of each panel.
The standard deviation is also shown in Figure~\ref{fig:mean2d} (bottom row)
at different times. We clearly notice the propagation of the variance
due to the parametric uncertainty with time
and along the tsunami's reflections. 

        
\subsubsection{Sensitivity analysis}
A global sensitivity analysis is performed to quantify the contribution of each
uncertain parameter to the variance in water surface elevation. To this end, we calculate 
the total sensitivity index using the PC coefficients as shown in Equation~\eqref{eq:T-hard}~\citep{Alexanderian2012,Sudret,Crestaux}. The evolution of the total sensitivity index
of each of the uncertain parameters is shown in Figure~\ref{fig:sens} at the four gauges. 
The Manning's $n$ coefficient at the shore $n_2$ is clearly dominant and contributes
the most to the variance in the water surface elevation compared to the other two 
Manning's $n$ coefficients $n_1$ and $n_3$; this is true for almost the entire simulation time
and at the four gauges. The Manning's $n$ coefficient in the deep ocean $n_{3}$ at gauge number 21419 exhibits a higher sensitivity index during the first hour of simulation however, as noted previously, none of the gauge locations see a significant variance between simulations until after the first hour of simulation.
The on-shore Manning's $n$ coefficient, $n_{1 }$, has no effect on the variance in
the water surface elevation for the entire simulation time at all gauges.

In 2D, the sensitivity analysis shows also that $n_2$ is dominant
at the same regions where we observe significant variance in water surface elevation. This is
indicated from Figure~\ref{fig:sens2d} that shows the total sensitivity index
for $n_1$ (top row) $n_2$ (center row) and $n_3$ (bottom row)
at different times.

\subsubsection{Response surface}
In addition to the statistical and sensitivity analysis, the PC surrogate 
can be used to construct a response surface for the uncertain input parameters.
This is achieved by sampling the PC surrogate for different values of the 
canonical vector of random variables  $\xxi$ within the prior
range. Since $n_1$ shows insignificant contribution to the 
uncertainty in the model output, we only consider the response surface
of $n_2$ versus $n_3$ for a fixed value of $n_1=0.035$ as illustrated in Figure~\ref{fig:response2}
at $t=7200~s$. The most striking features are the relatively flat
horizontal contours in the $n_3$ direction suggesting that water surface elevation depends
only mildly on $n_3$ even during peak tsunami events. This is true for gauges 21401, 21413 and 21418. However,
at gauge 21419, the horizontal contours are observed in the $n_2 $ direction
meaning that water surface elevation depends only mildly on $n_2$.